---
title: "Curtis Analysis"
output: html_document
---
```{r Preping_Dataframes}
library(tidyverse)
library(lubridate)

`%!in%` <- Negate(`%in%`)

## Reading in 2021 vegetation data
veg_2021 <- as_tibble(read.csv(
  "/Users/isaacbailey/Library/CloudStorage/OneDrive-UW-Madison/Research/Curtis_Prairie_DA/Curtis_Analysis/Final_2021_Curtis_Sample_04062022.csv", header = TRUE))
names(veg_2021)[1] <- "Plot"

# Reclassifying total cover into single integer
veg_2021$Total_Cover <- replace(veg_2021$Total_Cover, veg_2021$Total_Cover == "\"1-25\"", 12.5)
veg_2021$Total_Cover <- replace(veg_2021$Total_Cover, veg_2021$Total_Cover == "\"26-50\"", 37.5)
veg_2021$Total_Cover <- replace(veg_2021$Total_Cover, veg_2021$Total_Cover == "\"51-75\"", 62.5)
veg_2021$Total_Cover <- replace(veg_2021$Total_Cover, veg_2021$Total_Cover == "\"76-95\"", 85.5)
veg_2021$Total_Cover <- replace(veg_2021$Total_Cover, veg_2021$Total_Cover == "\"96-100\"", 98)
veg_2021$Total_Cover <- as.numeric(veg_2021$Total_Cover)

# Reclassifying species cover into single integer
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"<1\"", 0.5)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"1-5\"", 2.5)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"6-25\"", 15.5)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"26-50\"", 38)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"51-75\"", 63)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"76-95\"", 85.5)
veg_2021$Sp_Cover <- replace(veg_2021$Sp_Cover, veg_2021$Sp_Cover == "\"96-100\"", 98)
veg_2021$Sp_Cover <- as.numeric(veg_2021$Sp_Cover)

# Reclassifying dates
veg_2021$Date <- mdy(veg_2021$Date)

# Clean out canopy cover
veg_2021 %>% filter(!is.na(Canopy_Cover)) %>% filter(is.na(Sp_Cover)) -> just_canopy
veg_2021 %>% anti_join(just_canopy) -> veg_2021

## Reading in 2002 data
veg_snyder <- as_tibble(read.csv("/Users/isaacbailey/Library/CloudStorage/OneDrive-UW-Madison/Research/Curtis_Prairie_DA/Curtis_Analysis/IB_and_MC_Combo_11_10_21 - 2002_Data.csv", header = TRUE))

# Reading in environmental data
cross_walk <- as_tibble(read.csv("/Users/isaacbailey/Library/CloudStorage/OneDrive-UW-Madison/Research/Curtis_Prairie_DA/Curtis_Analysis/Snyder_crosswalk.csv", header = TRUE))

# creating snyder dataframe
ds <- data.frame(veg_snyder)
ds <- merge(ds, cross_walk, by = "X2021_Plot")

# creating 2021 dataframe
dmci <- data.frame(veg_2021$Plot, veg_2021$Scientific_name)
dmci <- merge(dmci, cross_walk, by.x = "veg_2021.Plot", by.y = "X2021_Plot")

# Creating dataframe without rare species

nr_dmci <- veg_2021
richness_by_spec <- nr_dmci %>% group_by(Scientific_name) %>% tally()
richness_by_plot <- nr_dmci %>% group_by(Plot) %>% tally()
nr_spec <- filter(richness_by_spec, n > 10)
nr_dmci <- filter(nr_dmci, Scientific_name %in% nr_spec$Scientific_name)

```


```{r variance_of_species_richness_and_kriging}

# For ecological analyses
library(vegan)

# Spatial data processing
library(sp)
library(sf)

# For the geostats
library(automap)
library(gstat)

# For mapping
library(patchwork)
library(viridis)
library(ggplot2)

# Setting up dataframes

spat_df <- as.tibble(cbind(dmci$veg_2021.Plot, dmci$Easting, dmci$Northing, dmci$NumSpec))
spat_df <- aggregate(cbind(V2, V3, V4) ~ V1, data = spat_df, FUN = mean)

ggplot(data = spat_df, aes_(x = spat_df$V2, y = spat_df$V3, color = spat_df$V4)) +
  geom_point(size = 2) +
  scale_fill_viridis(option = "A") +
  labs(title = "Species Richness Per Plot in 2002", x="Longitude", y = "Latitude")+
  theme_bw()

ggplot(data = spat_df, aes_(x = spat_df$V2, y = spat_df$V3, color = richness_by_plot$n)) +
  geom_point(size = 2) +
  scale_fill_viridis(option = "A") +
  labs(title = "Species Richness Per Plot in 2021", x="Longitude", y = "Latitude")+
  theme_bw()

spat_sf <- st_as_sf(spat_df, coords = c("V2", "V3"), crs = 3069) %>%
  cbind(st_coordinates(.))

spat_v_snyder <- variogram(V4 ~ 1, as(spat_sf, "Spatial")
)
spat_v_dmci <- variogram(
  richness_by_plot$n ~ 1,
  as(spat_sf, "Spatial")
)

spat_v.fit_snyder <- fit.variogram(spat_v_snyder, vgm(c("Exp", "Mat", "Sph")), fit.kappa = TRUE, fit.method = 7)
plot(variogramLine(vgm(17.68, "Mat", 72.11875, kappa = 1.6), 1000), type = "l")

spat_v.fit_dmci <- fit.variogram(spat_v_dmci, vgm(c("Exp", "Mat", "Sph")), fit.kappa = TRUE, fit.method = 7)
plot(variogramLine(vgm(11.89528, "Sph", 401.7893, kappa = 0.5), 1000), type = "l")

grd_1_sf <- spat_sf %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_make_grid(
    cellsize = c(5, 5),
    what = "centers"
  ) %>%
  st_as_sf() %>%
  cbind(., st_coordinates(.))

grd_1_sp <- as(grd_1_sf, "Spatial")
gridded(grd_1_sp) <- TRUE
grd_1_sp <- as(grd_1_sp, "SpatialPixels")

SK_rich_snyder <- krige(
  V4 ~ 1,
  as(spat_sf, "Spatial"),
  grd_1_sp,
  model = spat_v.fit_snyder
)

SK_rich_dmci <- krige(
  richness_by_plot$n ~ 1,
  as(spat_sf, "Spatial"),
  grd_1_sp,
  model = spat_v.fit_dmci
)

image(SK_rich_snyder)
image(SK_rich_dmci)


```
```{r building_spatial_matrices_and_modeling_jaccard_distances}
library(betapart)

# pulling in data from data frames and filtering missing plots
dmci %>% filter(veg_2021.Plot != 905) -> spat_dmci
veg_snyder %>% filter(X2021_Plot != 1004) -> spat_snyder

# plot numbers and species for 2021 data
dmci_plot <- as.factor(spat_dmci$veg_2021.Plot)
dmci_spec <- as.factor(spat_dmci$veg_2021.Scientific_name)

# plot numbers and species for 2002 data
snyder_plot <- as.factor(spat_snyder$X2021_Plot)
snyder_spec <- as.factor(spat_snyder$X2021_Taxon)

# creating site/species matrix
dmci_mat <- table(dmci_plot, dmci_spec)

# corrects for values greater than 1 in 2021 matrix
dmci_mat[dmci_mat > 1] <- 1

# cereating site/species matrix for 2002 data
snyder_mat <- table(snyder_plot, snyder_spec)

# corrects for values greater than 1 in 2002 matrix
snyder_mat[snyder_mat > 1] <- 1

# creates a list of species found both sampling years
comb_spec <- unique(append(dmci_spec, snyder_spec))

# creates a matrix to append 2002 species not found in 2021 to 2021 species/site matrix
subset(comb_spec, comb_spec %!in% dmci_spec) -> ni_dmci
addon_dmci <- matrix(0,length(unique(dmci_plot)),length(ni_dmci))
colnames(addon_dmci) <- (ni_dmci)
row.names(addon_dmci) <- unique(dmci_plot)

# appends addon matrix to 2021 matrix
dmci_mat <- cbind(dmci_mat, addon_dmci)

# creates a matrix to append 2021 species not found in 2002 to 2002 species/site matrix
subset(comb_spec, comb_spec %!in% snyder_spec) -> ni_snyder
addon_snyder <- matrix(0,length(unique(snyder_plot)),length(ni_snyder))
colnames(addon_snyder) <- (ni_snyder)
row.names(addon_snyder) <- unique(snyder_plot)

# appends addon to 2002 matrix
snyder_mat <- cbind(snyder_mat, addon_snyder)

# reorders columns in matrices so that they are the same for betapart
dmci_mat <- dmci_mat[,order(colnames(dmci_mat))]
snyder_mat <- snyder_mat[,order(colnames(snyder_mat))]

# turns both matrices into betapart objects
beta_dmci_mat <- betapart.core(dmci_mat)
beta_snyder_mat <- betapart.core(snyder_mat)

# runs jaccard index between both times
jaccard <- beta.temp(beta_dmci_mat,beta_snyder_mat, "jaccard")

jaccard_spat_df <- data.frame(cbind(jaccard$beta.jac,
aggregate(cbind(spat_dmci$Northing, spat_dmci$Easting), list(spat_dmci$veg_2021.Plot), FUN = mean)))

ggplot(data = jaccard_spat_df, aes_(x = jaccard_spat_df$V2, y = jaccard_spat_df$V1, color = jaccard_spat_df$jaccard.beta.jac)) +
  geom_point(size = 2) +
  scale_color_viridis(option = "H", direction = -1) +
  labs(title = "Jaccard Distance Between Years", x="Longitude", y = "Latitude")+
  theme_gray()



```
===================================================================================================================
This chunk trims a phylogenetic tree for the species in our sample from the tree of Wisconsin flora from Spalink, Daniel et al. (2019). Not listed in the tree are the edge lengths for each branch. These are stored, but not graphically represented. As it stands now, this tree excludes individuals not identified to species,and some  This will be used to calculate PD, NTI, and NRI later in the document.

```{r Phylogeny}

library(ape)

#sort into distinct species
keep_species <- distinct(veg_2021, Scientific_name)

#reformat of species list for tree pruning
keep_species$Scientific_name <- as.character(keep_species$Scientific_name)
keep <- gsub(" ", "_", keep_species$Scientific_name)

#reading in large phylogenetic tree
bigtree <- read.tree("/Users/isaacbailey/Library/CloudStorage/OneDrive-UW-Madison/Research/Curtis_Prairie_DA/GC.ultrametric_genera_constrained.tre")

#finding taxa to remove
setdiff(bigtree$tip.label, keep) -> remove_taxa

#tree pruning
pruned.tree <- drop.tip(bigtree, remove_taxa)

#List species in tree
pruned.tree$tip.label

#writing out tree
write.tree(pruned.tree, file = '/Users/isaacbailey/Library/CloudStorage/OneDrive-UW-Madison/Research/Curtis_Prairie_DA/Curtis_Analysis/curtis.tre')

#plot tree
plot.phylo(pruned.tree, align.tip.label = TRUE)

```

===================================================================================================================


```{r}

```
```{r}

```
